### 2017-06-26 docs/manuscript/main.tex
Adding manuscript.

\documentclass[12pt]{article}
%
\usepackage{abstract,amsmath,amssymb,latexsym}
\usepackage{tikz-cd}
\usepackage{enumitem,epsf}
\usepackage{fullpage,tikz,float}
\usepackage[numbers]{natbib}
\usepackage[pdftex,colorlinks]{hyperref}

% locally defined macros
\usepackage{macros}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Use the following for revealing TODOs and appendices
% Options are: \draftfalse or \drafttrue
\newif\ifdraft
\draftfalse

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{
 \begin{minipage}[c]{1.05\textwidth}
 	\centerline{Towards a Continuous Hyperparameter}
 	\centerline{Representation for Neural Networks}
 \end{minipage}
}

\author{
	\vspace{1cm}
	William Guss\thanks{Email: wguss@berkeley.edu} \and
	Other Author\thanks{Email: other@berkeley.edu} \and 
	Other Author\thanks{Email: other@berkeley.edu}
}

\begin{document}

\maketitle
\thispagestyle{empty}

%% \vspace{-4cm}
%% \vspace*{\fill} {
%% \renewcommand{\abstractnamefont}{\normalfont\large\bfseries}
%% \renewcommand{\abstracttextfont}{\normalfont\normalfont}
%% \renewcommand{\baselinestretch}{1.0}
\begin{abstract}
%% \vspace{-0.25cm}
% \input{abstract}
\end{abstract}
%% }
%% \vspace*{\fill}

\ifdraft
\newpage
\input{todo}
\fi

\newpage
\thispagestyle{empty}

{\large
\topskip0pt
\vspace*{\fill}
\tableofcontents
\vspace*{\fill}
}

\newpage

\setcounter{page}{1}

% TODO: After planning phase move to sections.

\section{Planning \& Unorganized Results (*)}
\subsection{Motivation \& Goal}

\subsection{Questions/Hypothesis}

\subsection{Theory}

\subsubsection{Desired Results}

\subsubsection{Some Exposition}

\subsection{Experiments}

The following are a set of desired experiments to verify the newly proposed hyperparameter representation.

\subsection{Reading List}

\subsection{Related Notes}

\begin{itemize}
	\item \href{https://onedrive.live.com/edit.aspx/Documents/Machine%20Learning?cid=c05fef92a4e2f5cd&id=documents&wd=target%28Continuous%20Parameterization.one%7C577F6089-A9C8-40F1-A37C-474241ED5D20%2F%29
onenote:https://d.docs.live.net/c05fef92a4e2f5cd/Documents/Machine%20Learning/Continuous%20Parameterization.one#section-id={577F6089-A9C8-40F1-A37C-474241ED5D20}&end}{Continuous Hidden Dimension}
	\item \textbf{Some Thoughts on Local Search on Hidden Units.} Let $\scriptn$ be the $\mathfrak{n}$-discrete instatntiation of the following DFM
	\begin{equation*}
	\begin{tikzcd}
		\scripto: \boxed{\mathbb{R}^n} \arrow{r}{\mathfrak{d}} & \boxed{L^1(E(\gamma))} \arrow{r}{\mathfrak{f}} & \boxed{\mathbb{R}}
	\end{tikzcd}
	\end{equation*}
	where $E: \mathbb{R} \to \scriptl(\mathbb{R})$ is a function which parameterizes the domain over which the $\mathfrak{f}$-functional integrates.

	It was concluded in the last note that if $E(\gamma) = [0, \gamma] \in \scriptl(\mathbb{R})$ then we have the following problem for the piecewise constant
	parameterization of weights on $\mathfrak{f}, \mathfrak{d}.$ Let $F: \mathbb{R} \to \mathbb{R}$ be some loss function, and then computation of the local gradient ascent path
	gives
	\begin{equation*}
	\begin{aligned}
		\frac{\partial F}{\partial \gamma} &= \frac{d F}{d y^2} \frac{\partial y^2}{\partial \gamma} \\
		 &= \frac{d F}{d y^2} \cdot \left[\frac{\partial }{\partial \gamma}  \int_{[0, \gamma]} \sum_{k=1}^\infty  [\sigma \circ \mathfrak{d}(x)](u)  \chi_{k\cdot [0,1]}(u) W^1_{k}\ d\mu(u)   \right]_\mathfrak{n} \\
		 &= \frac{d F}{d y^2} \cdot \left[ \sum_{k=1}^\infty  [\sigma \circ \mathfrak{d}(x)](\gamma)  \chi_{k\cdot [0,1]}(\gamma) W^1_{k}    \right]_\mathfrak{n} \\
		 &= \frac{d F}{d y^2} \cdot  y^1_{\floor*{\gamma}} W^1_{\floor*{\gamma}}.
	\end{aligned}
	\end{equation*}

	In otherwords, gradient ascent on $F$ with respect to $\gamma$ will increase $\gamma$ if the error will decrease when the contribution of the last output neuron is increased (in magnitude); that is, if $\gamma' > \gamma$ then $(\gamma - \floor*{\gamma})$ increases, and thus $E$ decreases by virtue of the term 
	\begin{equation*}
		\int_{\floor*{\gamma}\cdot [0,1]} y^1(u) W^1_{\floor*{\gamma}}\ d\mu(u) = (\gamma - \floor*{\gamma}) y^1_{\floor*{\gamma}} W^1_{\floor*{\gamma}}
	\end{equation*} increasing. Searching over $\gamma$ is effecitvely the same as spending extra time changing the weight $W_{\floor*{\gamma}}^1$ using two linearly dependent parameters, $(\gamma - \floor*{\gamma})$ and $W_{\floor*{\gamma}}^1$, itself\footnote{An additonal conclusion is, at least by analogy, that local search on $E(\gamma)$ at any one place assumes that adjacent neurons have similar values }.

	  Thus we are led to the question: \emph{Is hyperparameter search a matter of model capacity or model accuracy, and in that distinction, does optimizing hyperparameters with respect to model accuracy coorespond to optimization on model capactiy and visa versa?}  Before we define more specifically model capacity in the context of this question, let us examine this question in two contexts.\

	   Above, we noted that a local search on $\gamma$ decreased error in exactly the same fashion as standard gradient descent, but a step in $\gamma$ of more than integral amount can increase error. To see this let $k = \floor*{\gamma}$. When $\Delta \gamma > 1$ then the $(k+1)$th neuron is then "enabled" so-to-speak. However, this $(k+1)$th neuron may perform a computation that increases error and so in the next step of gradient descent $\Delta \gamma$ would be negative, retreating away from the added model capacity of a randomly intiialized $(k+1)$th neuron. That is not to say that $\gamma$ might not increase again, repeating the process, or in the limit of such oscilations the update $ W^1_{k+1} -\alpha\partial E/\partial W^1_{k+1} \to W^1_{k+1}$, will eventually contribute to model accuracy, but relying on these dynamics as a result with no guarentees of convergence is questionable. Despite the fact that $\scriptn$ may need additional model capacity\footnote{There are functions which are unlearnable without a sufficient number of neurons for example.}, local search on capacity with respect to accuracy may not yield the required capacity to increase accuracy in the limit.

	   \todo{Include brief analysis of richard's paper.}

\end{itemize}

\subsection{Timeline}

T

% \input{related}
% \input{setting}
% \input{skeletons}
% \input{results}
% \input{compkerspa}
% \input{proofs}
% \input{discussion}

\ifdraft
\appendix
% \input{appendix}
\fi

% \subsection*{Acknowledgments}

\bibliographystyle{plainnat}
\bibliography{bib}

\end{document}